---
title: "Course Project of Practical Machine Learning"
author: "JohnnyChiu"
date: "2015/3/21"
output: html_document
---

###1. Load the data into r and take a quick look at our dataset. I also split our training data set into two groups for cross validation, called 'intrain_training' and 'intrain_testing'.
```{r, cache=TRUE}
setwd('/Users/JohnnyChiu/Desktop/檔案總管/學習/coursera/JohnHopkins-DataScience/Practical Machine Learning/project/')
training<-read.csv('pml-training.csv')
testing<-read.csv('pml-testing.csv')
dim(training)
dim(testing)
summary(training)

library(caret)
inTrain <- createDataPartition(y=training$classe,p=0.75, list=FALSE)
intrain_training <- training[inTrain,]
intrain_testing <- training[-inTrain,]
```

I will subset some of the elements out, and make every variable into type numeric for model fitting.
```{r}
library(dplyr)
intrain_training[intrain_training=='']<-NA
sub_intrain_training<-intrain_training[,!is.na(intrain_training[1,])]
sub_intrain_training<-sub_intrain_training[,c(-1,-2)]
dim(sub_intrain_training)

intrain_testing[intrain_testing=='']<-NA
sub_intrain_testing<-intrain_testing[,!is.na(intrain_testing[1,])]
sub_intrain_testing<-sub_intrain_testing[,c(-1,-2)]
dim(sub_intrain_testing)

for(i in 1:(dim(sub_intrain_training)[2]-1)){
  sub_intrain_training[,i]<-as.numeric(sub_intrain_training[,i])
}

for(i in 1:(dim(sub_intrain_testing)[2]-1)){
  sub_intrain_testing[,i]<-as.numeric(sub_intrain_testing[,i])
}

```

###2. Build models on the training dataset

***First, the rpart model***
```{r, cache=TRUE}

modFit_rpart <- train(classe ~ .,method="rpart",data=sub_intrain_training)
print(modFit_rpart$finalModel)
library(rattle)
fancyRpartPlot(modFit_rpart$finalModel)
predict(modFit_rpart,newdata=sub_intrain_testing)
confusionMatrix(sub_intrain_testing$classe,predict(modFit_rpart,newdata=sub_intrain_testing))
#table(sub_intrain_testing$classe)
#table(predict(modFit_rpart,newdata=sub_intrain_testing))
```



***Second, the naive bayes model***
```{r,cache=TRUE}
modFit_nb <- train(classe ~ ., data=sub_intrain_training,method="nb")
predict(modFit_nb,sub_intrain_testing)
table(predict(modFit_nb,sub_intrain_testing),sub_intrain_testing$classe)
confusionMatrix(sub_intrain_testing$classe,predict(modFit_nb,newdata=sub_intrain_testing))
```

***Third, the Linear discriminant analysis(lda) model***
```{r,cache=TRUE}
modFit_lda <- train(classe ~ ., data=sub_intrain_training,method="lda")
predict(modFit_lda,sub_intrain_testing)
table(predict(modFit_lda,sub_intrain_testing),sub_intrain_testing$classe)
confusionMatrix(sub_intrain_testing$classe,predict(modFit_lda,newdata=sub_intrain_testing))

```

***Fourth, the random forest model. However, it seems it will take very long time to run.***
```{r}
##method1
#it takes too ling to run it.
# modFit_ranforest <- train(classe~ .,data=sub_intrain_training,method="rf",prox=TRUE)
# modFit_ranforest
# getTree(modFit_ranforest$finalModel,k=2)
# pred <- predict(modFit,testing); testing$predRight <- pred==testing$Species
# table(pred,testing$Species)

##method2
library(randomForest)
set.seed(415)
modFit_ranforest2 <- randomForest(as.factor(classe) ~ ., data=sub_intrain_training, importance=TRUE, ntree=50)
varImpPlot(modFit_ranforest2)
predict(modFit_ranforest2, sub_intrain_testing)
confusionMatrix(sub_intrain_testing$classe,predict(modFit_ranforest2,newdata=sub_intrain_testing))
```
***The accuracy of each model***
* rpart:0.5326
* naive bayes:0.7606
* Linear discriminant analysis(lda):0.803
* Random Forest:0.9986

***Hence, we choose Random Forest to model our data.


###3. Choose the best model from the models that we built, and use the model we fit to predict our testing dataset  
```{r, cache=TRUE}
testing[testing=='']<-NA
testing<-testing[,!is.na(testing[1,])]
testing<-testing[,c(-1,-2)]
dim(testing)
for(i in 1:(dim(testing)[2]-1)){
  testing[,i]<-as.numeric(testing[,i])
}

predict(modFit_ranforest2, testing)
```

###4. to generate the file to submit to the coursera website.
```{r, cache=TRUE}
answers = c('B','A','B','A','A','E','D','B','A','A','B','C','B','A','E','E','A','B','B','B')

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```




